Concurrent Programming with CXL

### 2020-06-29 book/paper.tex
Formatting up to Chapter 19

The \texttt{synch} library supports a synchronized message queue,
similar to the \texttt{Queue} object in Python.
Its interface is as follows:
\begin{itemize}
\item \texttt{Queue()} returns a new message queue;
\item \texttt{enqueue(q, item)} adds \texttt{item} to the queue pointed to by \texttt{q};
\item \texttt{dequeue(q)} waits for and returns an item on the queue pointed to by \texttt{q}.
\end{itemize}

@@

Figure~\ref{fig:actor} shows a CXL solution to the
``multiple producer / multiple consumer problem'' (with an unbounded buffer)
using an actor (as illustrated in Figure~\ref{fig:actorpic}.
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \texttt{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \texttt{balance} and \texttt{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

The \texttt{pc\_actor} process has two arguments.  \texttt{q} is a pointer
to the queue on which it receives messages.  \texttt{nrequests} is the total
number of requests that it will handle.  We need the latter because in CXL
all processes are required to terminate.  In a more realistic implementation,
the \texttt{pc\_actor} would be in an infinite loop awaiting requests.

@@

The \emph{Alternating Bit Protocol} is suitable for our purposes.
We assume that there are two unreliable network channels: one from Alice
to Bob and one from Bob to Alice.
Alice and Bob each maintain a zero-initialized sequence number,
\texttt{s\_seq} and \textit{r\_seq} resp.
Alice sends a network message to Bob containing an
application message as \emph{payload} and Alice's sequence number
as \emph{header}.
When Bob receives such a network message, Bob returns an
an \emph{acknowledgment} to Alice, which is a network message
containing the sequence number in the message that Bob received.

@@

Because of the network's properties, we can model each
channel as a variable that either contains a network message or nothing
(we use \texttt{()} in the model).  Let \texttt{s\_chan} be the channel
from Alice to Bob and \textit{r\_chan} the channel from Bob to Alice.
\texttt{net\_send(pchan, m, reliable)} models sending a message \texttt{m} to
\texttt{\^{}pchan}, where \texttt{pchan} is either \texttt{\&s\_chan}
or \texttt{\&r\_chan}.
The method places either \texttt{m} (to model a successful send)
or \texttt{()} (to model loss) in \texttt{\^{}pchan}.
The use of the \texttt{reliable} flag will be explained later.
\texttt{net\_recv(pchan)} models checking \texttt{\^{}pchan} for
the next message.

Method \texttt{app\_send(m)} tries \emph{retransmitting} \texttt{m} until
an acknowledgment is received.
Method \texttt{app\_recv(reliable)} returns the next successfully received
message.
Figure~\ref{fig:abptest} shows how the methods may be used to send and receive
a stream of \texttt{NMSGS} messages reliably.
It has to be bounded, because model checking requires a finite model.

Only the last invocation of \texttt{app\_recv(reliable)} is invoked with
\texttt{reliable == True}.  It causes the last acknowledgment to be sent
reliably.  It allows the receiver (Bob) to stop, as well as the sender (Alice)
once the last acknowledgment has been received.
Without something like this, either the sender may be left hanging waiting
for the last acknowledgment, or the receiver waiting for the last message.

>>>>

The \texttt{synch} library supports a synchronized message queue,
similar to the \texttt{Queue} object in Python.
Its interface is as follows:
\begin{itemize}
\item \texttt{Queue()} returns a new message queue;
\item \texttt{enqueue}($q$, \textit{item}) adds \textit{item} to the queue pointed to by $q$;
\item \texttt{dequeue}($q$) waits for and returns an item on the queue pointed to by $q$.
\end{itemize}

@@

Figure~\ref{fig:actor} shows a CXL solution to the
``multiple producer / multiple consumer problem'' (with an unbounded buffer)
using an actor (as illustrated in Figure~\ref{fig:actorpic}.
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \textit{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \textit{balance} and \textit{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

The \texttt{pc\_actor} process has two arguments.  $q$ is a pointer
to the queue on which it receives messages.  \textit{nrequests} is the total
number of requests that it will handle.  We need the latter because in CXL
all processes are required to terminate.  In a more realistic implementation,
the \texttt{pc\_actor} would be in an infinite loop awaiting requests.

@@

The \emph{Alternating Bit Protocol} is suitable for our purposes.
We assume that there are two unreliable network channels: one from Alice
to Bob and one from Bob to Alice.
Alice and Bob each maintain a zero-initialized sequence number,
\textit{s\_seq} and \textit{r\_seq} resp.
Alice sends a network message to Bob containing an
application message as \emph{payload} and Alice's sequence number
as \emph{header}.
When Bob receives such a network message, Bob returns an
an \emph{acknowledgment} to Alice, which is a network message
containing the sequence number in the message that Bob received.

@@

Because of the network's properties, we can model each
channel as a variable that either contains a network message or nothing
(we use \texttt{()} in the model).  Let \textit{s\_chan} be the channel
from Alice to Bob and \textit{r\_chan} the channel from Bob to Alice.
\texttt{net\_send}(\textit{pchan}, $m$, \textit{reliable}) models sending
a message $m$ to
\texttt{\^{}}\textit{pchan}, where \textit{pchan} is either \&\textit{s\_chan}
or \&\textit{r\_chan}.
The method places either $m$ (to model a successful send)
or \texttt{()} (to model loss) in \texttt{\^{}}\textit{pchan}.
The use of the \textit{reliable} flag will be explained later.
\texttt{net\_recv}(\textit{pchan}) models checking \texttt{\^{}pchan} for
the next message.

Method \texttt{app\_send}($m$) tries \emph{retransmitting} $m$ until
an acknowledgment is received.
Method \texttt{app\_recv}(\textit{reliable}) returns the next successfully received
message.
Figure~\ref{fig:abptest} shows how the methods may be used to send and receive
a stream of \texttt{NMSGS} messages reliably.
It has to be bounded, because model checking requires a finite model.

Only the last invocation of \texttt{app\_recv}(\textit{reliable}) is invoked with
\textit{reliable} == \texttt{True}.  It causes the last acknowledgment to be sent
reliably.  It allows the receiver (Bob) to stop, as well as the sender (Alice)
once the last acknowledgment has been received.
Without something like this, either the sender may be left hanging waiting
for the last acknowledgment, or the receiver waiting for the last message.
